---
title: "Using UMI4Cats to analyze UMI-4C data"
author: "Mireia Ramos Rodríguez & Marc Subirana-Granes"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    fig_width: 6
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Using UMI4Cats to analyze UMI-4C data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
Hello stranger! If you are here, that means you've successfully completed the UMI-4C protocol and got some sequencing results.

```{r}
devtools::load_all('/imppc/labs/lplab/share/marc/repos/umi4cats')
```

# Quick start
```{r, eval=F}

wd <- system.file(package = "UMI4Cats")

raw_dir <- file.path(wd, "extdata/raw_dir")
wk_dir <- file.path(wd, "extdata/wk_dir")
threads <- 1
bait_seq <- 'ACCTAGAAGGATATGCGCTTGC'
bait_pad <- 'GCGTTAGA'
res_e <- "GATC"
cut_pos <- 0 
fastqmultx <- 'fastq-multx'
bowtie2 <- 'bowtie2'
ref_gen <- '/imppc/labs/lplab/share/marc/refgen/hg19/hg19.fa'
samtools <- 'samtools'
genomic_track <- file.path(wd, "extdata/genomicTracks/dpnII_genomicTrack")
py_umi4cats <- '/imppc/labs/lplab/share/env/general/bin/python3'


umi4CatsContacts(py_umi4cats,
                 raw_dir,
                 wk_dir,
                 threads,
                 bait_seq,
                 bait_pad,
                 res_e,
                 ref_gen,
                 genomic_track,
                 trimmomatic,
                 bowtie2,
                 fastqmultx,
                 samtools)

## Analyzing single profiles -------------------

ctrCounters <- list.files(file.path(wd, "extdata/wk_dir/rst"),
                          full.names = T)[1]


ctrl.proc <- processUMI4C(ctrCounters,
                          factor=0.005)

plotUMI4C(ctrl.proc,
          xlim = c(36870000, 37230000)) # Start and end coordintes to plot



```

```{r, eval=F}
## Comparison of two profiles ------------------
treat.proc <- processUMI4C(treatCounters,
                          factor=0.005) 

treatCounters <- list.files(file.path(wd, "extdata/wk_dir/rst"),
                          full.names = T)[2]

# 1) Get statistics on differential contacts
differentialContacts(umi4c_obj=treat.proc,
                     ref_umi4c_obj=ctrl.proc) # TODO: Add some promoters to test CURRENTLY NOT WORKING WITH THE EXAMPLE
library(UMI4Cats)
```


# Quick start
```{r, eval=F}
## Processing FASTQ files ----------------------
# TODO: Include quick set of functions for analyzing fastq files

## Analyzing single profiles -------------------
ctrl.proc <- processUMI4C("../data-raw/B0M6_CCR1-5_TFRC_enhUmiRecount.txt",
                          factor=0.01)

plotUMI4C(ctrl.proc,
          ymax=3000, # Maximum UMIs value to plot in y axis
          xlim=c()) # Start and end coordintes to plot

tret.proc <- processUMI4C("../data-raw/B3M6_CCR1-5_TFRC_enhUmiRecount.txt",
                          factor=0.01) # TODO: This file is currently missing!

```


# Processing UMI-4C data


# Analyzing UMI-4C data
Once the processing step is finished, you should end up with a text file containing the following informtion:
- chr bait
- start bait
- chr contact
- start contact
- number of UMIs

This file needs to be loaded into R and then processed

## Analysis of single profiles

The first function one need to use to analyze the processed UMI-4C results in `processUMI4C`. The input of this function can be either a character with the filename where the result of the processing is stored or an already loaded data.frame with that information. 

The input data.frame needs to have 5 columns:

- Bait coordinates (Columns 1:2): `chr` and `start`.
- Contact coordinates (Columns 3:4): `chr` and `start`.
- Number of UMIs supporting the contact (Column 5). 

The `processUMI4C` function will automatically obtain the bait coordinates from the input data.frame. If you want to provide a custom name for the bait that will show in the plot, you need to use the argument `bait_name`.

If for some reason you need to specify custom bait coordinates, you can do so by using the `bait_coordinates` argument. This argument should be a `GRanges` object with the coordinates of the custom bait and one `mcols` element representing the name for the bait (if you want to leave it empty, just add the mcols with value `""`).  



```{r, fig.width=5, fig.height=3}
ctrl.proc <- processUMI4C("../data-raw/B0M6_CCR1-5_TFRC_enhUmiRecount.txt",
                          factor=0.01)

plot <- plotUMI4C(ctrl.proc)
plot
```


- `processUMI4C`
- `plotUMI4C`

## Comparative analysis of profiles
- `processComparisonUMI4C`
- `plotComparisonUMI4C`
- `differentialContacts`
